\chapter{Conclusion}

A speech synthesizer system is built for Urdu using Festival speech synthesis system \cite{thefestivalspeechsynthesissystem} and and Festvox \cite{festvox}. We used recorded speech of 70 minutes which consists of over 700 sentences. It contains almost 10000 tokens and over 5500 unique words. This data is divided in testing and training data in 10:90 proportion and used for training of the model to be used in synthesize speech. We used lexicon of Hindi for Urdu in order to build this system. A text preprocessing unit is also developed which process any number, dates or time string in input text. 

The resulting system is tested by 47 native Urdu speakers and found to be intelligible but not very natural. The synthesized words having same sound but having different first or last character are also recognizable. The resulting voice is intelligible and usable but it not very close to natural voice. It is because system is developed using Hindi lexicon and letter to sound rules. Some character in Hindi are spoken differently as compared to Urdu. For example, \texturdu{‫غ‬} in Hindi is spoken as \texturdu{گ}. Apart from this, the size of training data also effects quality of synthesized speech. The quality of synthesized speech also depends on size of speech corpus used. In current development of the system, we did not consider context of word in input data which also effects pronunciation of the word. In future, we can use parts of speech tagger in order get correct pronunciation of each word. 



